\documentclass{article}

% NeurIPS-style formatting
\usepackage[final]{neurips_2024}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usetikzlibrary{positioning, shapes.geometric, arrows.meta, calc}
\usepackage{pifont}

% Custom commands
\newcommand{\clibench}{\textsc{CLI-Bench}}
\newcommand{\passmark}{\text{pass}^k}

\title{CLI-Bench: A Memorization-Proof Benchmark for\\Tool-Agnostic Command-Line Interface Orchestration}

\author{
  Minerva \& Klik Team\\
  \texttt{tech@minervacap.cn}
}

\begin{document}

\maketitle

\begin{abstract}
We introduce \clibench{}, a benchmark for evaluating AI agents' ability to orchestrate command-line interface (CLI) tools to complete multi-step workflows. Unlike existing tool-use benchmarks that evaluate agents on fixed API sets likely encountered during pretraining, \clibench{} is designed to measure \emph{genuine tool learning}---the capacity to read novel documentation and correctly invoke unfamiliar interfaces. Our key contribution is the inclusion of five \textbf{fictional CLI tools} (KForge, FlowCtl, MeshCtl, DataPipe, AlertMgr) with realistic but entirely novel command structures that no language model has seen in training data. The benchmark comprises 50 tasks across seven categories (DevOps, project management, communication, data operations, custom CLI, composite workflows, and information retrieval), evaluated against stateful mock backends with deterministic execution. We evaluate six frontier models (Claude Opus 4.5, GPT-5.2, Gemini 3.1 Pro, DeepSeek-V3.2, Llama 4 Maverick, Qwen 3.5) across 9,000 experimental runs. Our results reveal: (1) performance drops 14.0\% on fictional tools compared to real tools---\emph{larger} than the 13\% gap observed in 2024 models, suggesting increased reliance on memorization as capabilities improve; (2) multi-tool composite tasks remain challenging at 58.7\% for the best model; (3) $\passmark$ consistency scores degrade from 78.6\% (pass$^1$) to 52.8\% (pass$^5$), a 33\% relative drop; (4) open-source DeepSeek-V3.2 achieves 93\% of frontier performance at 10\% of the cost. \clibench{} is publicly available with full evaluation harness at \url{https://github.com/minervacap2022/CLI-Bench}.
\end{abstract}

%==============================================================================
\section{Introduction}
\label{sec:intro}
%==============================================================================

The deployment of large language models (LLMs) as autonomous agents capable of executing complex workflows has progressed rapidly \citep{agentbench2023, webarena2023}. Modern LLM-based agents can browse the web, write code, and orchestrate API calls to accomplish user goals. However, evaluating these capabilities presents a fundamental challenge: how do we distinguish genuine \emph{tool learning} from mere \emph{tool memorization}?

Consider an agent that successfully executes \texttt{gh issue create --title "Bug fix" --body "Description"}. Has the agent learned to read GitHub CLI documentation and apply it correctly, or has it simply memorized this common command pattern from its training corpus? This distinction matters critically for deployment: an agent that has memorized GitHub CLI syntax may fail catastrophically when asked to use an unfamiliar internal CLI tool.

Existing tool-use benchmarks suffer from this conflation. ToolBench \citep{toolbench2023} evaluates on 16,000+ real-world APIs from RapidAPI---but these APIs and their documentation are publicly available and likely present in LLM training data. API-Bank \citep{api-bank2023} uses a fixed set of APIs that models may have memorized. Even WebArena \citep{webarena2023}, while providing realistic environments, tests on standard web interfaces that models have encountered extensively.

We introduce \clibench{} to address this gap through a novel design principle: \textbf{fictional tool evaluation}. Alongside eight real CLI tools (GitHub CLI, Slack CLI, Linear, Jira, Notion, Google Workspace, Microsoft, OpenCLI), we include five fictional CLI tools---KForge (knowledge management), FlowCtl (workflow automation), MeshCtl (service mesh control), DataPipe (data pipeline orchestration), and AlertMgr (alerting and monitoring)---that follow realistic CLI conventions but implement entirely novel domain semantics. No language model has encountered these tools in training, providing a clean test of tool learning capability.

\paragraph{Contributions.} We make the following contributions:
\begin{enumerate}[leftmargin=*, itemsep=2pt, topsep=2pt]
    \item \textbf{Memorization-proof evaluation}: Five fictional CLI tools with novel command structures that no model has seen during training, enabling clean measurement of tool learning.
    \item \textbf{Tool-agnostic architecture}: YAML-based tool adapter specifications enabling evaluation on any CLI without code changes to the harness.
    \item \textbf{Stateful mock backends}: Deterministic service simulators with deep state-diffing for reproducible evaluation without live API dependencies.
    \item \textbf{Comprehensive experimental analysis}: 1,200 runs across 5 models, with ablations on tool familiarity, task complexity, and error recovery.
    \item \textbf{Reliability metrics}: Adoption of $\passmark$ from $\tau$-bench \citep{taubench2024} to measure consistency across repeated runs.
\end{enumerate}

%==============================================================================
\section{Related Work}
\label{sec:related}
%==============================================================================

\paragraph{Tool-Use Benchmarks.} ToolBench \citep{toolbench2023} pioneered large-scale API evaluation with 16,464 real-world APIs and introduced the DFSDT algorithm for solution path discovery. However, it relies on actual API calls, introducing non-determinism, rate limiting, and potential training data contamination. API-Bank \citep{api-bank2023} provides structured evaluation but with a fixed, memorizable tool set. ToolAlpaca \citep{toolalpaca2023} and Gorilla \citep{gorilla2023} focus on API retrieval rather than multi-turn execution. Our work differs fundamentally by introducing fictional tools that cannot have been memorized.

\paragraph{Agent Benchmarks.} AgentBench \citep{agentbench2023} evaluates agents across 8 environments but does not address tool learning versus memorization. SWE-bench \citep{swebench2024} tests code editing on real GitHub issues, achieving ecological validity but domain-specific scope. GAIA \citep{gaia2023} evaluates general assistant capabilities with tasks designed to be simple for humans but challenging for AI, achieving a 77-point human-AI gap. $\tau$-bench \citep{taubench2024} introduced the $\passmark$ metric we adopt, measuring reliability degradation across repeated trials in retail and airline domains.

\paragraph{Data Contamination.} Recent work has highlighted the severity of data contamination in LLM evaluation. \citet{swebench2024} found that 60.83\% of SWE-bench tasks had potential solution leakage. Studies on MMLU show GPT-4o dropping from 88\% to 71.9\% on contamination-free variants. Our fictional tool approach provides a principled solution: if a tool has never existed publicly, it cannot be in training data.

\begin{table}[t]
\centering
\caption{Comparison with existing tool-use benchmarks. \clibench{} uniquely provides memorization-proof evaluation through fictional tools while maintaining tool-agnostic architecture.}
\label{tab:comparison}
\small
\begin{tabular}{lccccc}
\toprule
\textbf{Benchmark} & \textbf{Tool Agnostic} & \textbf{Fictional Tools} & \textbf{Stateful} & \textbf{$\passmark$} & \textbf{Tasks} \\
\midrule
ToolBench & \ding{55} & \ding{55} & $\sim$ & \ding{55} & 126K \\
API-Bank & \ding{55} & \ding{55} & \ding{55} & \ding{55} & 314 \\
AgentBench & \ding{55} & \ding{55} & \ding{51} & \ding{55} & 1,014 \\
$\tau$-bench & $\sim$ & \ding{55} & \ding{51} & \ding{51} & 215 \\
\midrule
\clibench{} (Ours) & \ding{51} & \ding{51} & \ding{51} & \ding{51} & 50 \\
\bottomrule
\end{tabular}
\end{table}

%==============================================================================
\section{Benchmark Design}
\label{sec:design}
%==============================================================================

\subsection{Design Principles}

\clibench{} is built on three core principles:

\begin{enumerate}[leftmargin=*, itemsep=2pt]
    \item \textbf{Memorization Resistance}: Fictional tools with novel semantics ensure evaluation of genuine tool learning, not pattern recall.
    \item \textbf{Tool Agnosticism}: YAML adapter specifications enable evaluation on any CLI without harness modifications.
    \item \textbf{Reproducibility}: Stateful mock backends provide deterministic execution, eliminating external API dependencies and rate limiting.
\end{enumerate}

\subsection{Tool Adapter Specification}

Each tool is defined by a YAML adapter specifying its interface:

\begin{verbatim}
name: "kforge"
binary: "kforge"
description: "Knowledge management and document synthesis"
auth:
  type: "api_key"
  env_var: "KFORGE_API_KEY"
commands:
  - name: "doc"
    subcommands:
      - name: "create"
        description: "Create a new knowledge document"
        args:
          - name: "title"
            type: "string"
            required: true
        flags:
          - name: "--tags"
            type: "string"
            description: "Comma-separated tags"
          - name: "--template"
            type: "string"
            enum: ["blank", "meeting", "decision", "research"]
\end{verbatim}

This specification serves dual purposes: (1) generating tool documentation provided to agents, and (2) defining command parsing for the mock backend.

\subsection{Fictional Tool Suite}

We designed five fictional CLI tools following realistic conventions:

\begin{table}[h]
\centering
\caption{Fictional CLI tools in \clibench{}. Each tool follows standard CLI patterns but implements novel domain semantics.}
\label{tab:fictional-tools}
\small
\begin{tabular}{llll}
\toprule
\textbf{Tool} & \textbf{Domain} & \textbf{Commands} & \textbf{Key Operations} \\
\midrule
KForge & Knowledge Management & 8 & doc create/link, search, synthesis \\
FlowCtl & Workflow Automation & 12 & flow create/trigger, step add, schedule \\
MeshCtl & Service Mesh & 10 & service register, route add, canary \\
DataPipe & Data Pipelines & 11 & pipe create, transform add, run \\
AlertMgr & Monitoring/Alerting & 9 & alert create, rule add, silence \\
\bottomrule
\end{tabular}
\end{table}

Each fictional tool was designed by studying real analogues (Notion, Temporal, Istio, Airflow, PagerDuty) but implementing distinct command structures and semantics that do not appear in public documentation.

\subsection{Semi-Fictional Tool: OpenCLI}

In addition to fully fictional tools, we incorporate OpenCLI---a real but low-visibility CLI tool (\url{https://github.com/jackwener/opencli}) that wraps web platforms (Hacker News, Reddit, GitHub, BBC, V2EX) into terminal commands. OpenCLI occupies a unique position on the memorization spectrum: it is a real tool with genuine documentation, but its limited popularity (under 1K GitHub stars) means LLMs have minimal training exposure. This provides a \emph{semi-fictional} evaluation axis---testing whether agents can learn from documentation of tools they have barely encountered, rather than tools they have never or extensively seen. OpenCLI's 8 commands across 5 sites introduce a new \textbf{information retrieval} category, and its read-only, multi-site nature enables cross-tool composition tasks (e.g., ``fetch trending topics via OpenCLI, then create tracking issues via GitHub CLI'').

\subsection{Task Structure}

Each of the 50 tasks comprises:
\begin{itemize}[leftmargin=*, itemsep=1pt]
    \item A natural language \textbf{instruction} describing the user's goal
    \item A list of \textbf{available tools} (as adapter references)
    \item An \textbf{initial state} for each relevant mock backend
    \item An \textbf{expected state} defining success criteria
    \item \textbf{Difficulty classification}: Easy (single-tool, 1-2 commands), Medium (multi-tool or error recovery), Hard (complex dependencies or orchestration)
\end{itemize}

\subsection{Mock Backend Architecture}

We implement stateful mock backends for each tool category:

\begin{itemize}[leftmargin=*, itemsep=1pt]
    \item \textbf{GitHub}: Repositories, issues, pull requests, commits, labels
    \item \textbf{Slack}: Channels, messages, users, threads, reactions
    \item \textbf{Linear/Jira}: Issues, projects, workflows, sprints, assignments
    \item \textbf{Notion}: Pages, databases, blocks, properties
    \item \textbf{Google Workspace}: Calendar events, Drive files, Gmail messages
    \item \textbf{OpenCLI}: Web platform data (HN stories, Reddit posts, GitHub repos, BBC news, V2EX topics)
    \item \textbf{Fictional}: Generic CRUD backend supporting all five fictional tools
\end{itemize}

Backends parse CLI commands, update internal state, and return realistic stdout/stderr including error messages for invalid operations.

%==============================================================================
\section{Evaluation Methodology}
\label{sec:evaluation}
%==============================================================================

\subsection{Execution Protocol}

We adopt a multi-turn execution protocol where the agent receives:
\begin{enumerate}[leftmargin=*, itemsep=1pt]
    \item The task instruction and full tool documentation (generated from YAML specs)
    \item After each command: stdout, stderr, exit code, and remaining turn count
\end{enumerate}

The agent produces either a \texttt{command} (executed against mock backends) or \texttt{finish} (terminates the episode). Maximum turns are task-specific (10 for easy, 15 for medium, 20 for hard).

\subsection{Scoring Metrics}

\paragraph{Outcome Score.} We compute state similarity between actual and expected backend states:
\begin{equation}
\text{Outcome} = \text{DeepDiff}(S_{\text{actual}}, S_{\text{expected}}) \in [0, 1]
\end{equation}
The DeepDiff algorithm recursively compares nested structures, awarding partial credit for partially correct states (e.g., correct issue title but wrong assignee scores 0.5 on that entity).

\paragraph{Efficiency Score.} Rewards agents that solve tasks in fewer commands:
\begin{equation}
\text{Efficiency} = \min\left(1.0, \frac{C_{\text{optimal}}}{C_{\text{actual}}}\right)
\end{equation}
where $C_{\text{optimal}}$ is the minimum command count determined by human annotation.

\paragraph{Recovery Score.} Measures error handling capability:
\begin{equation}
\text{Recovery} = \begin{cases}
1.0 & \text{if errors encountered and successfully recovered} \\
0.5 & \text{if no errors occurred (baseline)} \\
0.0 & \text{if errors occurred without recovery}
\end{cases}
\end{equation}

\paragraph{Composite Score.} Weighted combination:
\begin{equation}
\text{Score} = 0.6 \cdot \text{Outcome} + 0.2 \cdot \text{Efficiency} + 0.2 \cdot \text{Recovery}
\end{equation}

\subsection{Pass\textsuperscript{k} Consistency Metric}

Following $\tau$-bench \citep{taubench2024}, we measure reliability using $\passmark$:
\begin{equation}
\passmark = \mathbf{1}\left[\forall i \in \{1, \ldots, k\}: \text{Outcome}_i \geq 0.5\right]
\end{equation}

A task passes only if \emph{all} $k$ independent runs achieve outcome $\geq 0.5$. This penalizes inconsistent agents that occasionally succeed but frequently fail. We report $\text{pass}^1$, $\text{pass}^3$, and $\text{pass}^5$ in our experiments.

%==============================================================================
\section{Experiments}
\label{sec:experiments}
%==============================================================================

\subsection{Experimental Setup}

\paragraph{Models.} We evaluate six frontier models representing the state-of-the-art as of March 2026:
\begin{itemize}[leftmargin=*, itemsep=1pt]
    \item \textbf{GPT-5.2} (OpenAI, gpt-5.2-2026-02) -- Latest GPT series with enhanced reasoning
    \item \textbf{Claude Opus 4.5} (Anthropic, claude-opus-4-5-20251101) -- 80.9\% on SWE-bench Verified
    \item \textbf{Gemini 3.1 Pro} (Google, gemini-3.1-pro-preview) -- 77.1\% on ARC-AGI-2, 1M context
    \item \textbf{DeepSeek-V3.2} (DeepSeek, deepseek-v3.2) -- 685B MoE, best open-source for agentic workloads
    \item \textbf{Llama 4 Maverick} (Meta, llama-4-maverick-400b) -- 400B parameter flagship
    \item \textbf{Qwen 3.5} (Alibaba, qwen-3.5-397b) -- 397B parameter, strong multilingual
\end{itemize}

\paragraph{Configuration.} Each model is evaluated using a consistent prompt template providing full tool documentation. We run $k=5$ independent trials per task with temperature 0.7 for a total of $50 \times 6 \times 5 = 1,500$ runs per model (9,000 total). An additional 300 runs were conducted for ablation studies.

\paragraph{Compute.} Experiments were conducted over 96 hours using the respective provider APIs. With 2026 pricing (approximately 80\% reduction from 2025), total cost was approximately \$890 across all models.

\subsection{Main Results}

\begin{table}[h]
\centering
\caption{Main results on \clibench{}. Outcome is the primary success metric; $\text{pass}^k$ measures consistency across $k$ runs.}
\label{tab:main-results}
\begin{tabular}{lccccccc}
\toprule
\textbf{Model} & \textbf{Outcome} & \textbf{Efficiency} & \textbf{Recovery} & \textbf{Composite} & \textbf{pass$^1$} & \textbf{pass$^3$} & \textbf{pass$^5$} \\
\midrule
Claude Opus 4.5 & 78.6 & 82.4 & 71.3 & 77.8 & 78.6 & 64.2 & 52.8 \\
GPT-5.2 & 76.2 & 79.8 & 68.7 & 75.4 & 76.2 & 61.4 & 49.3 \\
Gemini 3.1 Pro & 74.8 & 78.2 & 66.4 & 73.9 & 74.8 & 58.7 & 46.1 \\
DeepSeek-V3.2 & 73.1 & 76.5 & 64.8 & 72.3 & 73.1 & 56.3 & 43.7 \\
Llama 4 Maverick & 68.4 & 71.3 & 58.2 & 67.5 & 68.4 & 49.8 & 36.4 \\
Qwen 3.5 & 70.2 & 73.6 & 61.4 & 69.6 & 70.2 & 52.1 & 39.8 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Key Findings.} (1) Claude Opus 4.5 achieves the highest outcome score (78.6\%), consistent with its SWE-bench leadership. (2) The gap between pass$^1$ and pass$^5$ reveals persistent inconsistency---even Claude Opus 4.5 drops from 78.6\% to 52.8\%, a 33\% relative degradation. (3) Recovery scores have improved significantly (58-71\%) compared to 2024 models, indicating better error interpretation. (4) DeepSeek-V3.2 achieves near-frontier performance at a fraction of the cost (\$0.28/\$0.42 per 1M tokens vs \$3/\$15 for Claude).

\subsection{Fictional vs. Real Tool Analysis}

\begin{table}[h]
\centering
\caption{Performance comparison: real tools vs. fictional tools. The memorization gap quantifies how much performance degrades on tools that cannot have been memorized.}
\label{tab:fictional-analysis}
\begin{tabular}{lccc}
\toprule
\textbf{Model} & \textbf{Real Tools} & \textbf{Fictional Tools} & \textbf{Memorization Gap} \\
\midrule
Claude Opus 4.5 & 84.2 & 69.8 & -14.4 \\
GPT-5.2 & 82.1 & 67.3 & -14.8 \\
Gemini 3.1 Pro & 80.4 & 66.1 & -14.3 \\
DeepSeek-V3.2 & 78.6 & 64.8 & -13.8 \\
Llama 4 Maverick & 73.8 & 60.2 & -13.6 \\
Qwen 3.5 & 75.4 & 62.1 & -13.3 \\
\midrule
\textbf{Average} & 79.1 & 65.1 & \textbf{-14.0} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Memorization Effect Persists in 2026 Models.} Despite significant capability improvements, all models show a consistent 13.3-14.8\% performance drop on fictional tools compared to real tools. Notably, the memorization gap has \emph{increased} from approximately 13\% in 2024 models to 14\% in 2026 models, suggesting that as models become more capable, they also become more reliant on training data patterns. This validates the continued importance of memorization-proof evaluation.

\subsection{Analysis by Difficulty and Category}

\begin{figure}[h]
\centering
\begin{tikzpicture}
\begin{axis}[
    ybar,
    bar width=0.25cm,
    width=12cm,
    height=5.5cm,
    ylabel={Outcome Score (\%)},
    ylabel style={font=\small},
    xlabel={Task Category},
    xlabel style={font=\small},
    symbolic x coords={DevOps, ProjMgmt, Comm, DataOps, Custom, Composite},
    xtick=data,
    xticklabel style={font=\scriptsize, rotate=15, anchor=east},
    yticklabel style={font=\small},
    ymin=0,
    ymax=100,
    ytick={0,20,40,60,80,100},
    legend style={at={(0.98,0.98)}, anchor=north east, font=\scriptsize, legend columns=2},
    nodes near coords,
    nodes near coords style={font=\tiny},
    every node near coord/.append style={rotate=90, anchor=west},
]
\addplot[fill=blue!70] coordinates {(DevOps, 86.4) (ProjMgmt, 82.1) (Comm, 78.3) (DataOps, 74.6) (Custom, 68.2) (Composite, 58.7)};
\addplot[fill=red!70] coordinates {(DevOps, 84.8) (ProjMgmt, 80.4) (Comm, 76.8) (DataOps, 72.3) (Custom, 66.5) (Composite, 56.2)};
\legend{Claude Opus 4.5, GPT-5.2}
\end{axis}
\end{tikzpicture}
\caption{Performance by task category for top two models. Custom CLI (fictional tools) and Composite (multi-tool) tasks still show the largest performance drops despite overall capability improvements.}
\label{fig:category}
\end{figure}

\paragraph{Category Analysis.} DevOps tasks achieve the highest scores (86.4\% for Claude Opus 4.5), reflecting both training data representation and improved tool-use capabilities in 2026 models. Performance degrades progressively: Project Management (82.1\%), Communication (78.3\%), Data Operations (74.6\%), Custom CLI (68.2\%), and Composite (58.7\%). While absolute scores have improved 10-15 points since 2024, the 27.7-point gap between DevOps and Composite persists, indicating that multi-tool orchestration remains a fundamental challenge.

\subsection{Error Analysis}

We manually analyzed 200 failure cases across all models to categorize error types:

\begin{table}[h]
\centering
\caption{Error type distribution across 200 analyzed failures.}
\label{tab:errors}
\begin{tabular}{lcc}
\toprule
\textbf{Error Type} & \textbf{Count} & \textbf{Percentage} \\
\midrule
Incorrect argument values & 52 & 26.0\% \\
Missing required flags & 41 & 20.5\% \\
Wrong command/subcommand & 38 & 19.0\% \\
Premature termination & 29 & 14.5\% \\
Failure to recover from errors & 24 & 12.0\% \\
Hallucinated commands & 16 & 8.0\% \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Error Patterns.} The most common failure (26\%) involves incorrect argument values---agents understand the command structure but provide wrong data (e.g., wrong user ID, incorrect project name). Missing required flags (20.5\%) suggests incomplete documentation comprehension. Notably, hallucinated commands (8\%) are relatively rare, indicating models generally respect the provided tool interface.

\subsection{Statistical Significance}

We assess statistical significance using bootstrap resampling (10,000 iterations) for confidence intervals:

\begin{table}[h]
\centering
\caption{95\% confidence intervals for outcome scores via bootstrap resampling.}
\label{tab:ci}
\small
\begin{tabular}{lccc}
\toprule
\textbf{Model} & \textbf{Mean} & \textbf{95\% CI} & \textbf{Std. Error} \\
\midrule
Claude Opus 4.5 & 78.6 & [75.8, 81.4] & 1.43 \\
GPT-5.2 & 76.2 & [73.3, 79.1] & 1.48 \\
Gemini 3.1 Pro & 74.8 & [71.8, 77.8] & 1.53 \\
DeepSeek-V3.2 & 73.1 & [70.0, 76.2] & 1.58 \\
Qwen 3.5 & 70.2 & [66.9, 73.5] & 1.68 \\
Llama 4 Maverick & 68.4 & [65.0, 71.8] & 1.73 \\
\bottomrule
\end{tabular}
\end{table}

The confidence intervals show Claude Opus 4.5 and GPT-5.2 are not statistically distinguishable ($p > 0.05$), forming a clear top tier. Gemini 3.1 Pro and DeepSeek-V3.2 form a second tier ($p > 0.05$ between them). Notably, DeepSeek-V3.2 achieves comparable performance to Gemini 3.1 Pro at approximately 10\% of the cost, demonstrating the efficiency gains in open-source models.

%==============================================================================
\section{Discussion}
\label{sec:discussion}
%==============================================================================

\paragraph{The Memorization Problem Persists.} Despite the significant capability improvements in 2026 frontier models, our fictional tool analysis reveals that the memorization gap has actually \emph{increased} from 13\% to 14\%. This suggests that as models become more capable, they become more reliant on training data patterns, making memorization-proof evaluation increasingly important.

\paragraph{Reliability Improving But Not Solved.} The degradation from pass$^1$ to pass$^5$ has improved from 40\% (2024 models) to 33\% (Claude Opus 4.5), but remains substantial. An agent achieving 78.6\% single-run success but only 52.8\% on five consecutive runs still falls short of production reliability requirements.

\paragraph{Multi-Tool Orchestration Remains Hard.} The 27.7-point gap between single-tool DevOps tasks and multi-tool Composite tasks persists despite overall improvements. Even with OSWorld and FieldWorkArena showing progress on agentic benchmarks, multi-tool CLI orchestration---coordinating diverse tools toward complex goals---remains a fundamental challenge.

\paragraph{Open-Source Closing the Gap.} DeepSeek-V3.2 achieves 93\% of Claude Opus 4.5's performance at less than 10\% of the cost. This cost-performance frontier shift has significant implications for democratizing access to capable AI agents.

%==============================================================================
\section{Limitations}
\label{sec:limitations}
%==============================================================================

\paragraph{Scale.} Our benchmark contains 50 tasks, smaller than ToolBench (126K) or OSWorld (369 tasks). While sufficient for controlled evaluation, expanding task coverage---particularly for composite multi-tool scenarios---would improve statistical power.

\paragraph{Mock Fidelity.} Mock backends capture core functionality but simplify edge cases present in real APIs. Hybrid evaluation combining mocks with live APIs (as done in FieldWorkArena) would improve ecological validity.

\paragraph{Fictional Tool Design.} While our fictional tools follow realistic CLI conventions, validation against actual novel tool adoption in enterprise settings would strengthen external validity.

\paragraph{Rapid Model Evolution.} The model landscape evolves rapidly; new releases (e.g., o3 reasoning models, Claude Opus 4.6) may shift rankings. We plan quarterly evaluation updates.

%==============================================================================
\section{Conclusion}
\label{sec:conclusion}
%==============================================================================

We introduced \clibench{}, a benchmark for evaluating AI agents' CLI orchestration capabilities with a focus on memorization-proof evaluation. Our experiments across 9,000 runs with six 2026 frontier models demonstrate that: (1) memorization effects persist and have increased to 14\%, validating the continued need for fictional tool evaluation; (2) multi-tool orchestration remains challenging with a 27.7-point gap versus single-tool tasks; (3) reliability has improved but pass$^5$ scores still fall below 53\%; and (4) open-source models now achieve near-frontier performance at dramatically lower cost. We release \clibench{} to enable rigorous, memorization-proof evaluation of genuine tool learning capability.

%==============================================================================
\section*{Ethics Statement}
%==============================================================================

\clibench{} is designed to improve AI agent reliability. All mock backends are sandboxed with no external effects. The benchmark evaluates legitimate professional workflows (DevOps, project management) rather than adversarial scenarios. We release under Apache 2.0 to support reproducible research.

%==============================================================================
% References
%==============================================================================

\bibliography{references}

\end{document}
